---
title: "Analyses of the `{palmerpenguins}` Data"
subtitle: "CIS 351 Sample Final Project"
author: "Anonymous"
format: 
  revealjs:
    incremental: true
---

```{r}
#| echo: false

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(ggpubr)

library(tinytable)
```

## 1. Introduction

Tell me something about your data.

## 2. Summary Stats

We examine the following numerical and categorical variables.

* **Numerical variables**: `body_mass_g`, `flipper_length_mm`, `bill_length_mm`
* **Categorical variables**: `species`, `island`, `sex`

**Note:** The histograms and bar charts included in this section are optional.

---

### 2.1. `body_mass_g`

```{r}
#| echo: false 

penguins |> 
  filter(!is.na(body_mass_g)) |> 
  summarize(
    min_bm = min(body_mass_g),
    median_bm = median(body_mass_g),
    mean_bm = round(mean(body_mass_g), 1),
    max_bm = max(body_mass_g),
    sd_bm = round(sd(body_mass_g),1)
  ) |> 
  tt()
```

```{r}
#| echo: false

penguins |> 
  filter(!is.na(body_mass_g)) |> 
  ggplot(
    aes(
      x = body_mass_g
    )
  ) +
  geom_histogram(
    fill = 'skyblue',
    color = 'blue'
  ) + 
  theme_bw() +
  labs(
    title = 'Distribution of Bill Length',
    x = 'Bill Length (mm)',
    y = 'Count'
  )
```

---

### 2.2. `flipper_length_mm` 

```{r}
penguins |> 
  filter(!is.na(flipper_length_mm)) |> 
  summarize(
    min_fl = min(flipper_length_mm),
    median_fl = median(flipper_length_mm),
    mean_fl = round(mean(flipper_length_mm)),
    max_fl = max(flipper_length_mm),
    sd_fl = round(sd(flipper_length_mm), 1)
  ) |> 
  tt()

penguins |> 
  filter(!is.na(flipper_length_mm)) |> 
  ggplot(
    aes(
      x = flipper_length_mm
    )
  ) +
  geom_histogram(
    fill = 'skyblue',
    color = 'blue'
  ) + 
  theme_bw() +
  labs(
    title = 'Distribution of Flipper Length',
    x = 'Flipper Length (mm)',
    y = 'Count'
  )
```

---

### 2.3. `bill_length_mm`

```{r}
penguins |> 
  filter(!is.na(bill_length_mm)) |> 
  summarize(
    min_bl = min(bill_length_mm),
    median_bl = median(bill_length_mm),
    mean_bl = round(mean(bill_length_mm), 1),
    max_bl = max(bill_length_mm),
    sd_bl = round(sd(bill_length_mm), 1)
  ) |> 
  tt()

penguins |> 
  filter(!is.na(bill_length_mm)) |> 
  ggplot(
    aes(
      x = bill_length_mm
    )
  ) +
  geom_histogram(
    fill = 'skyblue',
    color = 'blue'
  ) + 
  theme_bw() +
  labs(
    title = 'Distribution of Bill Length',
    x = 'Bill Length (mm)',
    y = 'Count'
  )
```

---

### 2.4. `species`

:::: {.columns}

::: {.column width="40%"}
```{r}
#| echo: false

penguins |> 
  filter(!is.na(species)) |> 
  summarize(
    Count = n(),
    .by = species
  ) |> 
  arrange(species) |> 
  tt()
```
:::

::: {.column width="60%"}
```{r}
#| echo: false

penguins |> 
  filter(!is.na(species)) |> 
  ggplot(
    aes(
      x = species,
      fill = species
    )
  ) +
  geom_bar() +
  scale_fill_brewer(palette = 'Dark2') +
  theme_bw() +
  labs(
    title = 'Count by Species',
    x = 'Species',
    y = 'Count'
  )
```
:::

::::

---

### 2.5. `island`

:::: {.columns}

::: {.column width="40%"}
```{r}
#| echo: false

penguins |> 
  filter(!is.na(island)) |> 
  summarize(
    Count = n(),
    .by = island
  ) |>
  arrange(island) |> 
  tt()
```
:::

::: {.column width="60%"}
```{r}
#| echo: false

penguins |> 
  filter(!is.na(island)) |> 
  ggplot(
    aes(
      x = island,
      fill = island
    )
  ) +
  geom_bar() +
  theme_bw() +
  scale_fill_brewer(palette = 'Dark2') +
  labs(
    title = 'Count by Island',
    x = 'Island',
    y = 'Count'
  )
```
:::

::::

---

### 2.6. `sex`

:::: {.columns}

::: {.column width="40%"}
```{r}
#| echo: false

penguins |> 
  filter(!is.na(sex)) |> 
  summarize(
    Count = n(),
    .by = sex
  ) |> 
  arrange(sex) |> 
  tt()
```
:::

::: {.column width="60%"}
```{r}
#| echo: false

penguins |> 
  filter(!is.na(sex)) |> 
  ggplot(
    aes(
      x = sex,
      fill = sex
    )
  ) +
  geom_bar() +
  scale_fill_brewer(palette = 'Dark2') +
  theme_bw() +
  labs(
    title = 'Count by Gender',
    x = 'Gender',
    y = 'Count'
  )
```
:::

::::

## 3. Data Visualizations

We will explore the relationships among the features of the `penguins` using the following visualizations.

* A scatter plot of `body_mass_g` and `flipper_length_mm`
* A boxplot of `body_mass_g` by `species`
* A stacked bar chart of `sex` and `size`

---

### 3.1. Body Mass vs Flipper Length

```{r}
#| echo: false

penguins |> 
  filter(!is.na(body_mass_g) & !is.na(flipper_length_mm)) |> 
  ggscatter(
    x = 'body_mass_g',
    y = 'flipper_length_mm',
    add = 'reg.line',
    add.params = list(
      color = 'blue'
    )  
  ) +
  stat_cor(
    label.x = 3000,
    label.y = 225
  ) +
  stat_regline_equation(
    label.x = 3000,
    label.y = 230
  ) +
  labs(
    title = 'Scatter Plot of Body Mass vs Flipper Length',
    x = 'Body Mass (g)',
    y = 'Flipper Length (mm)'
  )
```

---

### 3.2. Body Mass by Species

```{r}
#| echo: false

penguins |> 
  filter(!is.na(body_mass_g) & !is.na(species)) |> 
  ggplot(
    aes(
      x = species,
      y = body_mass_g,
      fill = species
    )
  ) +
  geom_boxplot() +
  theme_bw() +
  scale_fill_brewer(palette = 'Dark2') +
  stat_summary(
    fun = mean,
    fun.min = mean,
    fun.max = mean,
    geom = 'errorbar',
    color = 'red',
    width = 0.75,
    linewidth = 1.25,
    alpha = 0.5
  ) +
  labs(
    title = 'Plot of Body Mass by Species',
    x = 'Species',
    y = 'Body Mass (g)'
  ) +
  theme(legend.position = 'none')
```

---

### 3.3. Gender vs Size (Count)

```{r}
#| echo: false

penguins |> 
  filter(!is.na(body_mass_g) & !is.na(sex)) |> 
  mutate(
    size = case_when(
      body_mass_g >= mean(body_mass_g) ~ 'large',
      TRUE ~ 'small'
    )
  ) |> 
  ggplot(
    aes(
      x = sex,
      fill = size
    )
  ) +
  geom_bar(color = 'black') +
  scale_fill_brewer(palette = 'Dark2') +
  theme_bw() +
  labs(
    title = 'Stacked Bar Chart of Gender and Size',
    x = 'Gender',
    y = 'Size'
  )
```

---

### 3.3. Gender vs Size (Proportion)

```{r}
#| echo: false

penguins |> 
filter(!is.na(body_mass_g) & !is.na(sex)) |> 
  mutate(
    size = case_when(
      body_mass_g >= mean(body_mass_g) ~ 'large',
      TRUE ~ 'small'
    )
  ) |> 
  ggplot(
    aes(
      x = sex,
      fill = size
    )
  ) +
  geom_bar(
    color = 'black',
    position = 'fill'
  ) +
  scale_fill_brewer(palette = 'Dark2') +
  theme_bw() +
  labs(
    title = 'Stacked Bar Chart of Gender and Size',
    x = 'Gender',
    y = 'Size'
  )
```

## 4. Hypotheses Tests

### 4.1. ANOVA on `body_mass_g` by `species`

The hypotheses are

* $H_0$: the mean body mass are the same across all species
* $H_A$: at least one species has a mean body mass that differs from the others

---

### 4.1. ANOVA on `body_mass_g` by `species`
```{r}
#| echo: false

oneway.test(
  body_mass_g ~ species,
  data = penguins,
  var.equal = TRUE
)
```

Since the $p$-value is $p < 2.2 \times 10^{-16} < 0.05$, we reject $H_0$ and conclude that at least one species has a mean body mass that differs from others.

---

### 4.2. $\chi^2$-Test of Independence on `sex` and `size`

We define a new categorical `size` based on `body_mass_g`. 

* `size` is *large* if `body_mass_g` is greater than or equal to the mean `body_mass_g`.
* `size` is *small* if `body_mass_g` is less than the mean `body_mass_g`.

--- 

### 4.2. $\chi^2$-Test of Independence on `sex` and `size`

The hypotheses of the $\chi^2$-test are

* $H_0$: `sex` and `size` are independent variables
* $H_A$: `sex` and `size` are dependent variables

---

### 4.2. $\chi^2$-Test of Independence on `sex` and `size`

```{r}
penguins <- penguins |> 
  filter(!is.na(body_mass_g) & !is.na(sex)) |> 
  mutate(
    size = case_when(
      body_mass_g >= mean(body_mass_g) ~ 'large',
      TRUE ~ 'small'
    )
  )

chisq.test(
  table(
    penguins$sex,
    penguins$size
  )
)
```

Since the $p$-value is $p = 4.998 \times 10^{-5} < 0.05$, we reject $H_0$ and conclude that the variables `sex` and `size` are dependent.

## 5. Conclusion

Tell me what you have learned about your data.