---
title: "Week 4 Wednesday 02/12/2025"
author: "" 
format: 
  html:
    embed-resources: true
---


```{r}
#| message: false
# loading packges

library(tidyverse)

library(janitor)

library(palmerpenguins)
```

# Vectors (continued)

## Factors

Let us consider the vector `days` defined below and sort the vector using `sort()`.

```{r}
```

Yikes, not so good. `sort()` sorted the elements alphabetically. While this manner of sorting is preferred most of the time, there are lists that are not ordered alphabetically, like the weekdays.

We can overcome this problem by converting `days` into a factor. A factor is a character varaible with an extra property that you can impose how you want the levels to be ordered. The syntax of converting a (character) vector into a factor is `factor(x, levels =  c())`

```{r}
```

Also, note what is the difference (beside the ordering) in outputs between `sort(days)` and `sort(days_factor)`?

# Data Frames

A **data frame** is a structure that is a generalization of a vector. Whereas a vector is a one-dimensional object, a data frame can be a two-dimensional object. We use the function `data.frame()` to create a data frame from a collection of vectors.

```{r}
```

It is always a good idea to apply `class()` to a new type of object.

```{r}
```

We can use `str()` to get details of the data frame.

```{r}
```

There are three ways to extract column(s) from a data frame. 

1. the `[]` operator
2. the `$` operator
3. the `select()` function (part of `tidyverse`)

```{r}
# extract the `id` column from `df`
```

```{r}
# extract the 2nd column from `df`
```

```{r}
# extract the 2nd and 3rd columns from `df`
```

```{r}
# extract the `age` and `name` columns from `df`
```

```{r}
# extract the `age` column from `df`
```

```{r}
```

```{r}
```

Let's use a real data frame: the `penguins` data frame from `{palmerpenguins}` package.

```{r}
```

We often deal with very large data frames where it is not immediately clear what are the columns. To see the list of column names, we use `names()`.

```{r}
```

```{r}
# check variable type of the columns
```

Similar to `names()`, there are many R functions that will help us understand our data frames better. One such function is `unique()`. It allows us to see how many unique values there are in a column. For example, suppose we wish to know what are the species in the `species` column.

```{r}
```

**Exercise** How many distinct islands are there in the `penguins` data frame?

```{r}
```

Now that we know how to get to know our data frame, let's talk about summary values of the columns. 

**Question:** What is the mean body mass of `penguins`?

```{r}
```
The code returns `NA` because the `body_mass_g` has at least one `NA`. To remedy this, we add an **optional argument** in the `mean()` function that removes the `NA`s for us.

```{r}
```

**Exercise:** Find the standard deviation of `penguins`'s flipper length.

```{r}
```