---
title: "CIS 351 Week 9 Monday 03/31/2025"
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(ggpubr)
```

## Bar Plots with Error Bars

**Example 1.** Create a bar plot of penguins' mean body mass by species with error bars.

```{r}

```

## Interlude: Why Data Visualizations?

Data visualizations serve a higher purpose than just pretty charts and diagrams. They provide deeper insights to our data that summary values do not tell us.

### Problems with relying on summary values

**Example 2.** Import the animals weights data using the snippet provided in the code chunk. Compute the mean weights of each group of animals, then plot the weights by group.

```{r}
#| message: false
 
animals <- read.csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/animal_weights.csv")

```

```{r}

```

The following is a more extreme example, but it hits harder on the importance of data visualization. Examine the data `datasaurus` from the course github page and find the mean and standard deviation of the $x$-and $y$-values for each group.

```{r}
#| message: false
 
datasauras <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/datasauras.csv")

```

```{r}

```


```{r datasets, fig.height=12, fig.width=9}

```

## Scatter plots

Scatter plots are typically used to illustrate relationship between two continuous variables.

**Example 3.** Create a scatter plot of penguins flipper length vs body mass.

```{r}

```

**Example 4.** Repeat Example 3, except coloring the points by species.

```{r}

```

**Example 5.** Create a grouped scatter plot, where each species has its own plot.

```{r}

```