---
title: "Week 3 Friday 02/07/2025"
author: "" 
format: 
  html:
    embed-resources: true
---


```{r}
#| message: false
# loading packges

library(tidyverse)

library(janitor)
```

# Vectors (continued)

## Working with vectors

Rather than working with individual elements in a vector, R allows us to work with all elements simultaneously.

```{r}
temp_SA <- c(56, 58, 63, 69, 75, 80, 82, 82, 80, 73, 66, 59)

temp_NY <- c(33, 35, 42, 53, 63, 72, 77, 76, 69, 58, 48, 38)
```

We can perform arithmetic on the entire vector by performing the same operating to the variable name.


```{r}
```


```{r}
```

We can also perform operations involving two vectors, provided that they have the same number of elements.


```{r}
```


```{r}
```

:::{.callout-tip title = 'R Function'}
The function `length()` tells us the number of elements of a vector.
:::

```{r}
```

## Vectorized logical operations

**Question:** Which elements of `temp_NY` are greater than 55?

```{r}
```

**Exercise:** Which elements of `temp_SA` are not equal to 82?

```{r}
```

## The `%in%` operator

The `%in%` operator lets us determine whether the first object contains elements in a vector.

**Question:** Are there any elements in `temp_SA` that are equal to 66 or 75?
```{r}
```

**Exercise:** Are there any elements in `temp_NY` that are _not_ equal to 53 or 77?

```{r}
```

## Summary functions for vectors

:::{.callout-tip title = "R Function"}
* `sum()`: adds up all the elements in a (numeric) vector

* `mean()`: finds the mean value of a numeric vector

* `median()`: finds the median value of a numeric vector

* `sd()`: finds the standard deviation of a numeric vector

* `var()`: finds the variance of a numeric vector

* `min():` finds the minimum value in a vector

* `max():` finds the maximum value in a vector
:::

```{r}
```

Internally, R handles `TRUE` and `FALSE` as 1 and 0, respectively.

```{r}
```

This line of reasoning allows us to perform some summary operations.

**Question:** How many elements in `temp_NY` are less than 40?

```{r}
```

**Question:** How many elements in `temp_SA` are equal to 80 or 82?

```{r}
```

**Exercise:** What is the proportion of elements in `temp_NY` that are less than 40?

```{r}
```

## Defining integer sequences

To create a vector of sequence of consecutive integers, we use the `:` notation. 

```{r}
```

More generally, if we want to start the sequence with an integer $m$ and end with another integer $n$, then we use the syntax `m:n`. The `:` notation dispenses the need for `c()`.


```{r}
```

## Named vectors

We can assign names to each element of a vector using `names(df)`.

```{r}
```

## Extracting elements from a vector

R uses the square bracket, `[]`, notation to locate elements in a vector.


```{r}
# print the first element in `temp_SA`
```

```{r}
# print the 2nd through the 8th elements in `temp_SA`
```

```{r}
# print all but the first element in `temp_SA`
```

```{r}
# print all but the 4th element in `temp_SA`
```


```{r}
# print the 1st, 9th, 10th, 11th, and 12th elements in `temp_SA`
```

We can also use names of the elements instead of the element index (position).

```{r}
```

```{r}
```