---
title: "Week 3 Wednesday 02/05/2025"
author: "" 
format: 
  html:
    embed-resources: true
editor_options:
  chunk_output_type: console
---

# Types, continued

## `NA` values

A missing value, `NA`, is a special type of object. Like its logical cousins `TRUE` and `FALSE`, `NA` must be in all caps.

For example,
```{r}
NA
```

In general, when a value is missing in your data frame, it will be displayed as `NA` in R.

## Asking questions with logical operations

Let us define two new variables `x` and `y` by

```{r}
x <- 20*exp(1)

y <- 19*pi
```

**Question 1:** Is `x` equal to 40?

```{r}
x == 40
```

**Quesetion 2:** Is `y` less than or equal to 40?

```{r}
y <= 40
```

**Question 3:** Is `x` greater than `y`?

```{r}
x > y
```

**Question 4:** Is `x` not equal to `y`?

```{r}
x != y
```

In general, we use the following notations for logical operations

* **Equal to:** `==`
* **Not equal to:** `!=`
* **Greater than:** `>`
* **Less than:** `<`
* **Greater than or equal to:** `>=`
* **Less than or equal to:** `<=`

# Vectors

A **vector** is a collection of values (not necessarily of the same type). For thinking purpose, you can think of a vector is a *list* of objects. We use the `c()` function (concatenate or combine) to create a vector of values.

::: {.column-margin}
`list()` is a actually a different type of object in R.
:::

