---
title: "Week 5 Friday 02/21/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false
# loading packges

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(gapminder)
```

## Adding and modifying columns using `mutate()` 

`mutate()` allows us to redefine or create new columns based on existing columns in your dataframe.

**Example 1.** In `penguins` dataframe, convert the flipper length to inches.
```{r}

```

**Exercise 1.** In `penguins` dataframe, convert the body mass to pounds.

```{r}

```

A *very* useful function to go with `mutate()` is `case_when()`. `case_when()` gives you the ability to define values based on logical operations.

**Example 2.** In `gapminder` dataframe, filter out all countries in the Americas and label them by North, Central, or South America.
```{r}

```

**Exericse 2.** In `starwars` dataframe, convert the the heights of characters to feet and create a new column named `height_2` where a character whose height is greater than 180 cm is 'tall', between 160 and 180 is 'medium', less than 160 is short.

```{r}

```