---
title: "Week 6 Monday 02/24/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false
# loading packges

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(gapminder)
```

## Summarizing data using `summarize()`

We have encountered summary functions such as `mean()` and `sd()` while discussing vectors. One can similarly extend the use of summary functions to data frames via the `summarize()` function. More specifically, we can find summary statistics of specific column(s) in a data frame without needing to subset those columns.

**Example 1.** Find the mean height of all characters in `starwars`.

```{r}
```

### Use `is.na()` to detect `NA`s and count the number of `NA`s in a column.

When we pass a column through the `is.na()` function, R returns `TRUE` is an entry is `NA`, `FALSE` if it is not `NA`. Futhermore, since R processes `TRUE` and `FALSE` as 1 and 0, respectively, we can capitalize this by using `sum()` to get the number of `NA`s in a column.

**Example 1.2.** Determine whether `starwars$mass` containes any `NA`, if so, how many `NA`s are in the column.

```{r}
```

**Example 2.** Find the median and the maximum flipper lengths in `penguins`.

```{r}

```

**Exericse 1.** Find the mean and median ozone  of `airquality` dataset.

```{r}

```

### Summarizing data by groups

More often than not, we are interested in summary data of groups within a dataset. This will allow us to compare data between different groups.

**Example 3.** Find the mean body mass of `penguins` by species.

```{r}
```

**Example 4.** Find the mean ozone of `airquality` by month.

```{r}

```

**Exercise 2.** Find the mean GDP per capita of the Americas in `gapminder`.

```{r}

```