---
title: "Week 6 Monday 02/24/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false
# loading packages

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(gapminder)
```

## Summarizing data using `summarize()`

We have encountered summary functions such as `mean()` and `sd()` while discussing vectors. One can similarly extend the use of summary functions to data frames via the `summarize()` function. More specifically, we can find summary statistics of specific column(s) in a data frame without needing to subset those columns.

**Example 1.** Find the mean height of all characters in `starwars`.

```{r}
names(starwars)

starwars_mean_height <- starwars |> 
  summarize(
    mean_height = round(mean(height, na.rm = TRUE), 2)
  )

starwars_mean_height
```

### Use `is.na()` to detect `NA`s and count the number of `NA`s in a column.

When we pass a column through the `is.na()` function, R returns `TRUE` is an entry is `NA`, `FALSE` if it is not `NA`. Futhermore, since R processes `TRUE` and `FALSE` as 1 and 0, respectively, we can capitalize this by using `sum()` to get the number of `NA`s in a column.

**Example 1.2.** Determine whether `starwars$mass` containes any `NA`, if so, how many `NA`s are in the column.

Note that by passing the `mass` column to `is.na()`, R returns a vector of `TRUE` and `FALSE` values.
```{r}
is.na(starwars$mass)
```
If we further pass `is.na(starwars$mass)` to `sum()`, then R will count the number of `NA`s in the `mass` column.
```{r}
sum(is.na(starwars$mass))
```
In general, the following snippet of R code count the number of `NA` values in a specified column, assuming that the correct names for the data frame and column are provided.

```{r}
count_NAs <- starwars |> 
  summarize(
    NA_count = case_when(
      sum(is.na(mass)) == 0 ~ 'No NAs',
      TRUE ~ paste(as.character(sum(is.na(mass))), 'NAs')
    )
  )
```
A quick explanation on the second case of `case_when()`. As shown above, `sum(is.na(mass))` is the number of `NA` values in `mass` column. Recall from earlier in the semester that `as.character()` is a function that converts its argument to characters. Thus `as.character(sum(is.na(mass)))` counts the number of `NA`s then converts the value to a character value. Lastly, the `paste()` function is used to "glue" multiple character strings together. In this instance, the two strings are `as.character(sum(is.na(mass)))` and 'NAs'.

**Example 2.** Find the median and the maximum flipper lengths in `penguins`.

```{r}
names(penguins)

flipper_length_sum <- penguins |> 
  summarize(
    median_FL = round(median(flipper_length_mm, na.rm = TRUE), 2),
    max_FL = max(flipper_length_mm, na.rm = TRUE)
  )

flipper_length_sum
```

**Exercise 1.** Find the mean and median ozone  of `airquality` dataset.

```{r}
names(airquality)

# column names are not in lower snake form; pass it to `clean_names()`
df_airquality <- airquality |> 
  clean_names()

ozone_sum <- df_airquality |> 
  summarize(
    mean_ozone = round(mean(ozone, na.rm = TRUE), 2),
    median_ozone = round(median(ozone, na.rm = TRUE), 2)
  )

ozone_sum
```

*Question:* Based on the mean and median ozone levels, can you make an educated guess about the shape of the ozone level distribution?

### Summarizing data by groups

More often than not, we are interested in summary data of groups within a dataset. This will allow us to compare data between different groups.

**Example 3.** Find the mean body mass of `penguins` by species.

```{r}
mean_BM_species <- penguins |> 
  summarize(
    mean_BM = round(mean(body_mass_g, na.rm = TRUE), 2),
    .by = species
  )

mean_BM_species
```

**Example 4.** Find the mean ozone of `airquality` by month.

```{r}
mean_ozone_month <- df_airquality |> 
  summarize(
    mean_ozone = round(mean(ozone, na.rm = TRUE), 2),
    .by = month
  )

mean_ozone_month
```

**Exercise 2.** Find the mean GDP per capita for countries in the Americas using `gapminder`.

Before we can find the mean GDP per captia by country, we need to first subset out those countries in Americas.
```{r}
names(gapminder)

df_gapminder <- gapminder |> 
  clean_names()

mean_gdp <- df_gapminder |> 
  filter(continent == 'Americas') |> 
  summarize(
    mean_GDP = round(mean(gdp_percap, na.rm = TRUE), 2),
    .by = country
  )

mean_gdp
```
