---
title: "Week 7 Wednesday 03/05/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false
# loading packages

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(gapminder)
```

## Combining datasets 

We very rarely get all the data we need in a single data frame. Most often, the data you need come from multiple dataframes and you need to spend some time combining the datasets into a single data frame. 


```{r}
bats <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/Bats_data.csv")

geo <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/Geo_data.csv")

weather <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/Weather_vars.csv")
```

**Example 1.** Clean the above 3 dataframes. Select site, date, foraging, and activity from `bats`.

```{r}

```

**Example 2.** Add latitudinal and longitudinal information to `bats`.

```{r}

```

**Exercise 1.** Add mean temperature and mean humidity to the dataframe from Example 2.