---
title: "Week 5 Wednesday 02/19/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false
# loading packges

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(gapminder)
```

## `{dplyr}` functions

The `{dplyr}` package contains functions that allows us to work with data frames more easily. The main `{dplyr}` functions are

* `select()`: extract column(s) from your data frame
* `filter()`: filter rows of your data based on a condition/conditions
* `mutate()`: create new columns or modify exiting columns in your data frame
* `summarize()`: aggregate information in your columns

There are other `{dplyr}` functions we will introduce throughout the semester, but the above four functions are the most frequently used when working with data frames.

### `select()`

We have seen and used `select()` before. It allows us to extract columns we want to work with. To remove a column/columns, we can put a `-` symbol in front of the column name.

**Example 1.** Select the temperature, month, and day columns from `airquality` dataset. 

```{r}
```

**Example 2.** Remove the depth and stations columns from `quakes` dataset.

```{r}
```

### Rename columns

* `select()`: 

```{r}
```

* `rename()`

```{r}
```

### `filter()`

The `filter()` function lets you filter specific rows based on logical operation(s) on column(s).

**Example 3:** Subset out the data in `gapminder` where the country is Australia.

```{r}
```

**Example 4:** From `gapminder`, subset out the Australian data after 1980.

```{r}
```

**Exercise 1:** From `gapminder`, find the life expectancy data on the US and Egypt post 1990.

```{r}
```

**Exercise 2:** From `penguins`, find the body mass of the Chinstrap penguins from either Dream or Biscoe Island.

```{r}

```