---
title: "CIS 351 Week 10 Monday 04/07/2025"
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(palmerpenguins)
```

## $t$-Test

$t$-test is one of the most commonly used inferential test. There are three types of $t$-test: one-sample $t$-test, two-sample $t$-test, and paired $t$-test. We will focus on the two-sample $t$-test as it is the most frequently used out of the three.

Generally speaking, the two-sample $t$-test is determine whether there is a difference between the mean value between two different groups. 

The hypotheses for a two-sample $t$-test are
$$
H_0: \mu_1 = \mu_2
$$
$$
H_A: \mu_1 \ne \mu_2
$$
In the above, $\mu_1$ and $\mu_2$ are the means of groups 1 and 2. The inequality in $H_A$ indicates that this is a two-tailed test. We can also perform the $t$-test with $<$ or $>$ in $H_A$; these would turn the $t$-test into a one-tailed test.

**Example 1.** Use the $t$-test to determine whether there is a difference between the body mass of penguins by gender.

Let $\mu_M$ and $\mu_F$ denote the mean body mass of the male and female penguins, respectively. The hypotheses are
$$
H_0: 
$$
$$
H_A: 
$$
Next, we should visualize the boxplot for the body mass distribution by gender.

```{r}
```
To perform the $t$-test, we use the `t.test()` function.

```{r}
```

**Example 2.** From `airquality`, use the $t$-test to determine whether the mean ozone level for August is greater than the mean ozone level for May.

$$
H_0: \mu_M = \mu_A
$$
$$
H_A: \mu_M < \mu_A
$$

```{r}

```


```{r}

```