---
title: "CIS 351 Week 10 Friday 04/11/2025"
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(palmerpenguins)
```

## Chi-Square ($\chi^2$) Test of Independence

The $\chi^2$ test is a statistical test to determine whether there is a relationship between two categorical variables.

The null and alternative hypotheses of a $\chi^2$ test are.
* $H_0:$ the two categorical variables are independent, i.e., there is no relationship between them.
* $H_A:$ the two categorical variables are dependent, i.e., there is a relationship between them.

**Example 1.** From `penguins`, create a new variable `size`, which is defined by, if the body mass is greater than the mean body mass, then it is 'large'. If, on the other hand, a penguin's body mass is less than or equal to the mean body mass, then it is 'small'. Use the $\chi^2$ test to determine whether the variables `species` and `size` are dependent.

```{r}
penguins_1 <- penguins |> 
  filter(!is.na(body_mass_g)) |> 
  mutate(
    size = case_when(
      body_mass_g > mean(body_mass_g) ~ 'large',
      TRUE ~ 'small'
    )
  )

penguins_1 |> 
  ggplot(
    aes(
      x = species,
      fill = size
    )
  ) +
  geom_bar() +
  theme_bw() +
  labs(
    x = 'Species',
    y = 'Count'
  )

penguins_1 |> 
  ggplot(
    aes(
      x = species,
      fill = size
    )
  ) + 
  geom_bar(position = 'fill') +
  theme_bw() +
  labs(
    x = 'Species',
    y = 'Count'
  )
```

To perform the $\chi^2$ test in R, we use the `chisq.test()` function.

```{r}
test_1 <- chisq.test(
  table(
    penguins_1$species,
    penguins_1$size
  )
)

test_1
```

Since the $p$-value is $p < 2.2 \times 10^{-16} < 0.05$, we reject $H_0$ and conclude that `species` and `size` are dependent.