---
title: "Week 3 Wednesday 02/05/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

# Types, continued

## `NA` values

A missing value, `NA`, is a special type of object. Like its logical cousins `TRUE` and `FALSE`, `NA` must be in all caps.

For example,
```{r}
NA
```

In general, when a value is missing in your data frame, it will be displayed as `NA` in R.

## Asking questions with logical operations

Let us define two new variables `x` and `y` by

```{r}
```

**Question 1:** Is `x` equal to 40?

```{r}
```

**Quesetion 2:** Is `y` less than or equal to 40?

```{r}
```

**Question 3:** Is `x` greater than `y`?

```{r}
```

**Question 4:** Is `x` not equal to `y`?

```{r}
```

In general, we use the following notations for logical operations

* **Equal to:** `==`
* **Not equal to:** `!=`
* **Greater than:** `>`
* **Less than:** `<`
* **Greater than or equal to:** `>=`
* **Less than or equal to:** `<=`

# Vectors

A **vector** is a collection of values (not necessarily of the same type). For thinking purpose, you can think of a vector is a *list* of objects. We use the `c()` function (concatenate or combine) to create a vector of values.

::: {.column-margin}
`list()` is a actually a different type of object in R.
:::

**Example 1:** Create a vector consists the numbers 1, 10, 8, 44, pi, e. Check the data type of the vector.

```{r}
```

**Example 2:** Create a vector consists of 'Mon', 'Tue', 'Wed', 'Thu', 'Fri'. Check the data type of the vector.

```{r}
```

**Example 3:** Create a vector consists of 1, 44, 'apple', 'orange', pi. Check the data type of the vector.

```{r}
```

**Example 4:** Create a vector consists of 1, 5, TRUE, FALSE, -9, NA. Check the data type of the vector

```{r}
```

_**Remark:**_ In general, _vectors can only contain values of a single type._ The hierarchy of the data type is
 
 <p style='text-align:center'>
**Character > Numeric > Logical**
 </p>
