---
title: "Week 5 Friday 02/21/2025"
author: "" 
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false
# loading packges

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(gapminder)
```

## Summarizing data using `summarize()`

We have encountered summary functions such as `mean()` and `sd()` while discussing vectors. One can similarly extend the use of summary functions to data frames via the `summarize()` function. More specifically, we can find summary statistics of specific column(s) in a data frame without needing to subset those columns.

**Example 1.** Find the mean height of all characters in `starwars`.

```{r}
```

**Example 2.** Find the median and the maximum flipper lengths in `penguins`.

```{r}

```

**Exericse 1.** Find the mean and median ozone  of `airquality` dataset.

```{r}

```

### Summarizing data by groups

More often than not, we are interested in summary data of groups within a dataset. This will allow us to compare data between different groups.

**Example 3.** Find the mean body mass of `penguins` by species.

```{r}
```

**Example 4.** Find the mean ozone of `airquality` by month.

```{r}

```

**Exercise 2.** Find the mean GDP per capita of the Americas in `gapminder`.

```{r}

```