---
title: "CIS 351 Week 9 Monday 03/31/2025"
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(palmerpenguins)

library(ggpubr)
```

## Bar Plots with Error Bars

**Example 1.** Create a bar plot of penguins' mean body mass by species with error bars.

```{r}
plot_1 <- penguins |> 
  filter(!is.na(body_mass_g)) |> 
  summarize(
    mean_bm = mean(body_mass_g),
    sd_bm = sd(body_mass_g),
    .by = species
  ) |> 
  ggplot(
    aes(
      x = species,
      y = mean_bm
    )
  ) + 
  geom_bar(
    stat = 'identity', 
    fill = 'skyblue', 
    color = 'blue'
  ) +
  geom_errorbar(
    aes(
      ymin = mean_bm - sd_bm,
      ymax = mean_bm + sd_bm
    ),
    width = 0.4,
    color = 'orange',
    alpha = 0.9
  ) 

plot_1
```

## Interlude: Why Data Visualizations?

Data visualizations serve a higher purpose than just pretty charts and diagrams. They provide deeper insights to our data that summary values do not tell us.

### Problems with relying on summary values

**Example 2.** Import the animals weights data using the snippet provided in the code chunk. Compute the mean weights of each group of animals, then plot the weights by group.

```{r}
#| message: false
 
animals <- read.csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/animal_weights.csv")

weight_sum <- animals |> 
  summarize(
    mean_weight = round(mean(weight, na.rm = TRUE), 1),
    .by = group
  )

weight_sum
```

```{r}
plot_1 <- animals |> 
  ggplot(
    aes(
      x = group,
      y = weight
    )
  ) +
  geom_jitter(
    aes(color = group),
    alpha = 0.5
  ) + 
  stat_summary(
    geom = 'errorbar',
    fun = mean,
    fun.min = mean,
    fun.max = mean,
    width = 0.75,
    color = 'red'
  ) +
  theme_bw()

plot_1
```

The following is a more extreme example, but it hits harder on the importance of data visualization. Examine the data `datasaurus` from the course github page and find the mean and standard deviation of the $x$-and $y$-values for each group.

```{r}
#| message: false
 
datasauras <- read_csv("https://raw.githubusercontent.com/ckaoxvii/CIS351/refs/heads/main/datasauras.csv")

datasauras_x_sum <- datasauras |> 
  summarize(
    x_mean = round(mean(x, na.rm = TRUE), 2),
    x_sd = round(sd(x, na.rm = TRUE), 2),
    .by = group
  )

datasauras_x_sum
```

```{r}
datasauras_y_sum <- datasauras |> 
  summarize(
    y_mean = round(mean(y, na.rm = TRUE), 2),
    y_sd = round(sd(y, na.rm = TRUE), 2),
    .by = group
  )

datasauras_y_sum
```


```{r datasets, fig.height=12, fig.width=9}
plot_datasauras <- datasauras |> 
  ggplot(
    aes(
      x = x,
      y = y,
      color = group
    )
  ) +
  geom_point() +
  theme_void() +
  theme(legend.position = 'none') +
  facet_wrap(~dataset, ncol = 3)

plot_datasauras
```

## Scatter plots

Scatter plots are typically used to illustrate relationship between two continuous variables.

**Example 3.** Create a scatter plot of penguins flipper length vs body mass.

```{r}
plot_scatter <- penguins |> 
  filter(
    !is.na(flipper_length_mm),
    !is.na(body_mass_g)
  ) |> 
  ggscatter(
    x = 'flipper_length_mm',
    y = 'body_mass_g',
    add = 'reg.line',
    add.params = list(
      color = 'blue'
    ) 
  ) +
  stat_cor(
    label.x = 170,
    label.y = 6000
  ) +
  stat_regline_equation(
    label.x = 170,
    label.y = 5500
  )

plot_scatter
```

**Example 4.** Repeat Example 3, except coloring the points by species.

```{r}
plot_scatter_2 <- penguins |> 
  filter(
    !is.na(flipper_length_mm),
    !is.na(body_mass_g)
  ) |> 
  ggscatter(
    x = 'flipper_length_mm',
    y = 'body_mass_g',
    color = 'species',
    add = 'reg.line',
    add.params = list(
      color = 'black'
    ),
    alpha = 0.5
  ) +
  stat_cor(
    label.x = 170,
    label.y = 6000
  ) +
  stat_regline_equation(
    label.x = 170,
    label.y = 5500
  )

plot_scatter_2
```

**Example 5.** Create a grouped scatter plot, where each species has its own plot.

```{r}
plot_scatter_3 <- penguins |> 
  filter(
    !is.na(flipper_length_mm),
    !is.na(body_mass_g)
  ) |> 
  ggscatter(
    x = 'flipper_length_mm',
    y = 'body_mass_g',
    color = 'species',
    add = 'reg.line',
    add.params = list(
      color = 'black'
    ),
    alpha = 0.5
  ) +
  facet_wrap(~species) +
  stat_cor(
    label.y = 6000
  ) +
  stat_regline_equation(
    label.y = 5750
  )

plot_scatter_3
```