---
title: "CIS 351 Week 10 Monday 04/07/2025"
format: 
  html:
    embed-resources: true
---

```{r}
#| message: false

library(tidyverse)

library(janitor)

library(palmerpenguins)
```

## $t$-Test

$t$-test is one of the most commonly used inferential test. There are three types of $t$-test: one-sample $t$-test, two-sample $t$-test, and paired $t$-test. We will focus on the two-sample $t$-test as it is the most frequently used out of the three.

Generally speaking, the two-sample $t$-test is determine whether there is a difference between the mean value between two different groups.

The hypotheses for a two-sample $t$-test are $$
H_0: \mu_1 = \mu_2
$$ $$
H_A: \mu_1 \ne \mu_2
$$ In the above, $\mu_1$ and $\mu_2$ are the means of groups 1 and 2. The inequality in $H_A$ indicates that this is a two-tailed test. We can also perform the $t$-test with $<$ or $>$ in $H_A$; these would turn the $t$-test into a one-tailed test.

**Example 1.** Use the $t$-test to determine whether there is a difference between the body mass of penguins by gender.

Let $\mu_M$ and $\mu_F$ denote the mean body mass of the male and female penguins, respectively. The hypotheses are $$
H_0: \mu_M = \mu_F
$$ $$
H_A: \mu_M \ne \mu_F
$$ Next, we should visualize the boxplot for the body mass distribution by gender.

```{r}
penguins_1 <- penguins |> 
  filter(!is.na(sex))

penguins_1 |> 
  ggplot(
    aes(
      x = sex,
      y = body_mass_g,
      color = sex
    )
  ) + 
  geom_boxplot() +
  theme_bw() +
  stat_summary(
    fun = mean,
    fun.min = mean,
    fun.max = mean,
    geom = 'errorbar',
    color = 'black',
    width = 0.75,
    linewidth = 1.2
  ) +
  labs(
    title = 'Body Mass Distribution by Gender',
    x = 'Gender',
    y = 'Body Mass (g)'
  )
```

To perform the $t$-test, we use the `t.test()` function.

```{r}
t.test(
  body_mass_g ~ sex,
  data = penguins_1
)
```

Since the $p$-value is $p = 4.794 \times 10^{-16} < 0.05$, we reject the $H_0$ and conclude that the mean body mass for the penguins are different by gender.

**Example 2.** From `airquality`, perform the $t$-test to determine whether the mean ozone level for May is lower than the mean ozone level for August.

```{r}
airquality_1 <- airquality |> 
  clean_names() |> 
  filter(
    !is.na(ozone),
    month %in% c(5, 8)
  ) |> 
  mutate(
    month = case_when(
      month == 5 ~ 'May',
      month == 8 ~ 'August'
    )
  )

airquality_1 |> 
  ggplot(
    aes(
      x = month,
      y = ozone,
      color = month
    )
  ) +
  geom_boxplot() +
  theme_bw() +
  stat_summary(
    fun = mean,
    fun.min = mean,
    fun.max = mean,
    geom = 'errorbar',
    color = 'black',
    width = 0.75,
    linewidth = 1.2
  ) +
  labs(
    title = 'Ozone Levels in May and August',
    x = 'Month',
    y = 'Ozone Level'
  )
```

$$
H_0: \mu_M = \mu_A
$$

$$
H_A: \mu_M < \mu_A
$$

```{r}
t.test(
  ozone ~ month,
  data = airquality_1,
  alternative = 'greater' #alphabetically, August comes before May
)
```

With a $p$-value of $p = 0.0001084 < 0.05$, we reject $H_0$ and conclude that the mean ozone level for May is lower than the mean ozone level for August.
